---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# teal.modules.general

<!-- badges: start -->
<!-- badges: end -->

This package contains a set of standard `teal` modules for you to include in your `teal` applications. 
These include modules for:

- viewing data (including `tm_variable_browser`, `tm_data_table` and `tm_file_viewer`)
- visualizing data (`tm_g_scatterplot`, `tm_g_association` and many more)
- understanding missing and outlier values within your data (`tm_missing_data` and `tm_outliers`)
- performing simple data analysis (`tm_a_pca`, `tm_g_distribution` and `tm_a_regression`)

These modules work with both `CDISC` and general relational data.  

## Installation

This repository requires a personal access token to install see here [creating and using PAT](https://docs.github.com/en/github/authenticating-to-github/keeping-your-account-and-data-secure/creating-a-personal-access-token). Once this is set up, to install the latest released version of the package run:

```{r, eval=FALSE}
Sys.setenv(GITHUB_PAT = "your_access_token_here")
install.packages("devtools")
devtools::install_github("insightsengineering/teal.modules.general@*release")
```

<< Some comment about handling dependencies here >>

## Example

This is a basic example app showing a few modules from this package

```{r example}
library(teal.modules.general)

app <- init(
  data = teal_data(
    dataset("IRIS", x = iris)
  ),
  modules = modules(
    tm_data_table(),
    tm_variable_browser(),
    tm_t_crosstable(
      x = teal::data_extract_spec("IRIS",
        select = teal::select_spec(
          choices = c("Sepal.Length", "Sepal.Width"),
          multiple= TRUE
        )
      ),
      y = teal::data_extract_spec("IRIS",
        select = teal::select_spec(
          choices = "Species",
          multiple = FALSE,
          fixed = TRUE
        )
      )
    )
  )
)

```

Which is run using:
```{r, eval=FALSE}
runApp(app)
```


# Further Information

See package vignettes `browseVignettes(package = "teal.modules.general")` for examples using other modules from this package.
