---
title: "Decorate Module Output"
author: "NEST CoreDev"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Decorate Module Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(teal.modules.general)
```


## Introduction

The outputs produced by `teal` modules, like graphs or tables, are created by the module developer and look a certain way.
It is hard to design an output that will satisfy every possible user, so the form of the output should be considered a default value that can be customized.
Here we describe the concept of _decoration_, enabling the app developer to tailor outputs to their specific requirements without rewriting the original module code.

The decoration process is build upon transformation procedures, introduced in `teal`.
While `transformators` are meant to edit module's input, decorators are meant to adjust the module's output.
To distinguish the difference, modules in  `teal.modules.general` have 2 separate parameters:
`transformators` and `decorators`. 

To get a complete understanding refer the following vignettes:

- Transforming the input data in
[this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-input-data.html),
- Transforming module output in [this vignette](https://insightsengineering.github.io/teal/latest-tag/articles/transform-module-output.html).

## Outputs that can be decorated

It is important to note which output objects from a given module can be decorated.
The module function documentation's _Decorating Module_ section has this information.

You can also refer the table shown below to know which module outputs can be decorated.

| Module                 | Output (Class)                                      |
|------------------------|----------------------------------------------------|
| `tm_a_pca`            | elbow_plot (ggplot), circle_plot (ggplot), biplot (ggplot), eigenvector_plot (ggplot) |
| `tm_a_regression`     | plot (ggplot)                                      |
| `tm_g_association`    | plot (grob)                                        |
| `tm_g_bivariate`      | plot (ggplot)                                      |
| `tm_g_distribution`   | histogram_plot (ggplot), qq_plot (ggplot), summary_table (datatables), test_table (datatables) |
| `tm_g_response`       | plot (ggplot)                                      |
| `tm_g_scatterplot`    | plot (ggplot)                                      |
| `tm_g_scatterplotmatrix` | plot (trellis)                                  |
| `tm_missing_data`     | summary_plot (grob), combination_plot (grob), by_subject_plot (ggplot), table (datatables) |
| `tm_outliers`         | box_plot (ggplot), density_plot (ggplot), cumulative_plot (ggplot), table (datatables) |
| `tm_t_crosstable`     | table (ElementaryTable)                            |

Also, note that there are five different types of objects that can be decorated:
1. `ElementaryTable`
2. `ggplot`
3. `grob`
4. `datatables`
5. `trellis`

*Tip:* A general tip before trying to decorate the output from the module is to
copy the reproducible code and running them in a separate R session to
quickly iterate the decoration you want.

## Decorating `ElementaryTable`

Here's an example to showcase how you can edit an output of class `ElementaryTable`.
`rtables` modifiers like `rtables::insert_rrow` can be applied to modify this object.

```{r decorate_ElementaryTable, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  require(nestcolor)
  ADSL <- rADSL
})

insert_rrow_decorator <- function(default_caption = "I am a good new row") {
  teal_transform_module(
    label = "New row",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "new_row"), "New row", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                table <- rtables::insert_rrow(table, rtables::rrow(new_row))
              },
              new_row = input$new_row
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_t_crosstable(
      label = "Cross Table",
      x = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          choices = variable_choices(data[["ADSL"]], subset = function(data) {
            idx <- !vapply(data, inherits, logical(1), c("Date", "POSIXct", "POSIXlt"))
            return(names(data)[idx])
          }),
          selected = "COUNTRY",
          multiple = TRUE,
          ordered = TRUE
        )
      ),
      y = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          choices = variable_choices(data[["ADSL"]], subset = function(data) {
            idx <- vapply(data, is.factor, logical(1))
            return(names(data)[idx])
          }),
          selected = "SEX"
        )
      ),
      decorators = list(
        table = insert_rrow_decorator()
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_1, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_ElementaryTable")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `ggplot`

Here's an example to showcase how you can edit an output of class `ggplot`.
You can extend them using `ggplot2` functions.

```{r decorate_ggplot, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  require(nestcolor)
  ADSL <- rADSL
})

ggplot_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                plot <- plot + ggplot2::labs(caption = footnote)
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_a_regression(
      label = "Regression",
      response = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          label = "Select variable:",
          choices = "BMRKR1",
          selected = "BMRKR1",
          multiple = FALSE,
          fixed = TRUE
        )
      ),
      regressor = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          label = "Select variables:",
          choices = variable_choices(data[["ADSL"]], c("AGE", "SEX", "RACE")),
          selected = "AGE",
          multiple = TRUE,
          fixed = FALSE
        )
      ),
      decorators = list(
        plot = ggplot_caption_decorator("I am a Regression")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_2, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_ggplot")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `grob`

Here's an example to showcase how you can edit an output of class `grob`.
You can extend them using `grid` and `gridExtra` functions.

```{r decorate_grob, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  ADSL <- rADSL
})

grob_caption_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) {
      shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption)
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                footnote_grob <- grid::textGrob(
                  footnote,
                  x = 0, hjust = 0,
                  gp = grid::gpar(fontsize = 10, fontface = "italic", col = "gray50")
                )
                plot <- gridExtra::arrangeGrob(
                  plot,
                  footnote_grob,
                  ncol = 1,
                  heights = grid::unit.c(
                    grid::unit(1, "npc") - grid::unit(1, "lines"), grid::unit(1, "lines")
                  )
                )
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_association(
      ref = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          choices = variable_choices(
            data[["ADSL"]],
            c("SEX", "RACE", "COUNTRY", "ARM", "STRATA1", "STRATA2", "ITTFL", "BMRKR2")
          ),
          selected = "RACE"
        )
      ),
      vars = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          choices = variable_choices(
            data[["ADSL"]],
            c("SEX", "RACE", "COUNTRY", "ARM", "STRATA1", "STRATA2", "ITTFL", "BMRKR2")
          ),
          selected = "BMRKR2",
          multiple = TRUE
        )
      ),
      decorators = list(
        plot = grob_caption_decorator("I am a Association")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_3, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_grob")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `datatables`

Here's an example to showcase how you can edit an output of class `datatables`.
Please refer the [helper functions](https://rstudio.github.io/DT/functions.html) of the `DT` package to learn more about extending the `datatables` objects.
```{r decorate_datatables}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  require(nestcolor)
  ADSL <- rADSL
})
fact_vars_adsl <- names(Filter(isTRUE, sapply(data[["ADSL"]], is.factor)))
vars <- choices_selected(variable_choices(data[["ADSL"]], fact_vars_adsl))

dt_table_decorator <- function(color1 = "pink", color2 = "lightblue") {
  teal_transform_module(
    label = "Table color",
    ui = function(id) {
      selectInput(
        NS(id, "color"),
        "Table Color",
        choices = c("white", color1, color2),
        selected = "Default"
      )
    },
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |> within(
            {
              summary_table <- DT::formatStyle(
                summary_table,
                columns = attr(summary_table$x, "colnames")[-1],
                target = "row",
                backgroundColor = color
              )
            },
            color = input$color
          )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_distribution(
      dist_var = data_extract_spec(
        dataname = "ADSL",
        select = select_spec(
          choices = variable_choices(data[["ADSL"]], c("AGE", "BMRKR1")),
          selected = "BMRKR1",
          multiple = FALSE,
          fixed = FALSE
        )
      ),
      strata_var = data_extract_spec(
        dataname = "ADSL",
        filter = filter_spec(
          vars = vars,
          multiple = TRUE
        )
      ),
      group_var = data_extract_spec(
        dataname = "ADSL",
        filter = filter_spec(
          vars = vars,
          multiple = TRUE
        )
      ),
      decorators = list(
        summary_table = dt_table_decorator()
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_4, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_datatables")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```

## Decorating `trellis`

Here's an example to showcase how you can edit an output of class `trellis`.
`rtables` modifiers like `rtables::insert_rrow` can be applied to modify this object.

```{r decorate_trellis, message=FALSE}
library(teal.modules.general)

data <- teal_data(join_keys = default_cdisc_join_keys[c("ADSL", "ADRS")])
data <- within(data, {
  require(nestcolor)
  ADSL <- rADSL
  ADRS <- rADRS
})

trellis_subtitle_decorator <- function(default_caption = "I am a good decorator") {
  teal_transform_module(
    label = "Caption",
    ui = function(id) shiny::textInput(shiny::NS(id, "footnote"), "Footnote", value = default_caption),
    server = function(id, data) {
      moduleServer(id, function(input, output, session) {
        reactive({
          data() |>
            within(
              {
                plot <- update(plot, sub = footnote)
              },
              footnote = input$footnote
            )
        })
      })
    }
  )
}

app <- init(
  data = data,
  modules = modules(
    tm_g_scatterplotmatrix(
      label = "Scatterplot matrix",
      variables = list(
        data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(data[["ADSL"]]),
            selected = c("AGE", "RACE", "SEX"),
            multiple = TRUE,
            ordered = TRUE
          )
        ),
        data_extract_spec(
          dataname = "ADRS",
          filter = filter_spec(
            label = "Select endpoints:",
            vars = c("PARAMCD", "AVISIT"),
            choices = value_choices(data[["ADRS"]], c("PARAMCD", "AVISIT"), c("PARAM", "AVISIT")),
            selected = "INVET - END OF INDUCTION",
            multiple = TRUE
          ),
          select = select_spec(
            choices = variable_choices(data[["ADRS"]]),
            selected = c("AGE", "AVAL", "ADY"),
            multiple = TRUE,
            ordered = TRUE
          )
        )
      ),
      decorators = list(
        plot = trellis_subtitle_decorator("I am a Scatterplot matrix")
      )
    )
  )
)

if (interactive()) {
  shinyApp(app$ui, app$server)
}
```

```{r shinylive_iframe_5, echo = FALSE, out.width = '150%', out.extra = 'style = "position: relative; z-index:1"', eval = requireNamespace("roxy.shinylive", quietly = TRUE) && knitr::is_html_output() && identical(Sys.getenv("IN_PKGDOWN"), "true")}
code <- paste0(c(
  "interactive <- function() TRUE",
  knitr::knit_code$get("setup"),
  knitr::knit_code$get("decorate_trellis")
), collapse = "\n")

url <- roxy.shinylive::create_shinylive_url(code)
knitr::include_url(url, height = "800px")
```
