---
title: "Using cross table"
author: "Dawid Kałędkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Using cross table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Teal application to use cross table with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside an cross table module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(dplyr) # used to modify data sets
```

## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADSL2` A wide data set with subject data
3. `ADRS` A long data set with response data for subjects at different time points of the study
4. `ADTTE` A long data set with time to event data
5. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide", echo=2:6}
# nolint start
ADSL <- teal.modules.general::rADSL
ADSL2 <- teal.modules.general::rADSL %>%
  mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))
ADRS <- teal.modules.general::rADRS
ADTTE <- teal.modules.general::rADTTE
ADLB <- teal.modules.general::rADLB %>%
  mutate(CHGC = as.factor(case_when(
    CHG < 1 ~ "N",
    CHG > 1 ~ "P",
    TRUE ~ "-"
  )))
# nolint end
```

## Create an `app` variable

This is the most important section. We will use the [`teal::init`](https://insightsengineering.github.io/teal/reference/init) function to
create an app. The data will be handed over using [`teal.data::cdisc_data`](https://insightsengineering.github.io/teal.data/reference/cdisc_data). The app
itself will be constructed by multiple calls of `tm_t_crosstable` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- teal::init(
  data = teal.data::cdisc_data(
    teal.data::cdisc_dataset("ADSL", ADSL, code = "ADSL <- teal.modules.general::rADSL"),
    teal.data::cdisc_dataset(
      "ADSL2",
      ADSL2,
      keys = get_cdisc_keys("ADSL"),
      code = "ADSL2 <- teal.modules.general::rADSL %>%
              mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))"
    ),
    teal.data::cdisc_dataset("ADRS", ADRS, code = "ADRS <- teal.modules.general::rADRS"),
    teal.data::cdisc_dataset("ADTTE", ADTTE, code = "ADTTE <- teal.modules.general::rADTTE"),
    teal.data::cdisc_dataset("ADLB", ADLB,
      code = "ADLB <- teal.modules.general::rADLB %>%
            mutate(CHGC = as.factor(case_when(
            CHG < 1 ~ 'N',
            CHG > 1 ~ 'P',
            TRUE ~ '-'
            )))"
    ),
    check = TRUE
  ),
  modules = teal::modules(
    modules(
      label = "Cross table",
      tm_t_crosstable(
        label = "Single wide dataset",
        x = teal.transform::data_extract_spec(
          "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = names(ADSL)[5],
            multiple = TRUE,
            fixed = FALSE,
            ordered = TRUE
          )
        ),
        y = teal.transform::data_extract_spec(
          "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = names(ADSL)[6],
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),
      tm_t_crosstable(
        label = "Same long datasets (different subsets)",
        x = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = teal.transform::filter_spec(
            vars = "PARAMCD",
            choices = value_choices(ADLB, "PARAMCD", "PARAM"),
            selected = levels(ADLB$PARAMCD)[1],
            multiple = FALSE
          ),
          select = select_spec(
            choices = variable_choices(ADLB),
            selected = "AVISIT",
            multiple = TRUE,
            fixed = FALSE,
            ordered = TRUE,
            label = "Select variable:"
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = teal.transform::filter_spec(
            vars = "PARAMCD",
            choices = value_choices(ADLB, "PARAMCD", "PARAM"),
            selected = levels(ADLB$PARAMCD)[1],
            multiple = FALSE
          ),
          select = select_spec(
            choices = variable_choices(ADLB),
            selected = "LOQFL",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        )
      )
    )
  )
)
```

## Run the app

A simple `shiny::shinyApp` call will let you run the app.
Note that app is only displayed when running this code inside an R session.

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```
