---
title: "Using data table"
author: "coreDev team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Using data table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Teal application to use association plot with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside a data table module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
```

## Create data sets

Inside this app 3 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADTTE` A long data set with time to event data
3. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
# nolint start
data <- teal_data()
data <- within(data, {
  ADSL <- teal.modules.general::rADSL
  ADTTE <- teal.modules.general::rADTTE
  ADLB <- teal.modules.general::rADLB
})
datanames <- c("ADSL", "ADTTE", "ADLB")
datanames(data) <- datanames
join_keys(data) <- default_cdisc_join_keys[datanames]
# nolint end
```

## Create an `app` variable

This is the most important section. We will use the [`teal::init`](https://insightsengineering.github.io/teal/reference/init) function to
create an app. The data will be handed over using [`teal.data::cdisc_data`](https://insightsengineering.github.io/teal.data/reference/cdisc_data). The app itself will be constructed by multiple calls of `tm_data_table` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- teal::init(
  data = data,
  modules = teal::modules(
    # two-datasets example
    tm_data_table(
      label = "Two datasets",
      variables_selected = list(
        ADSL = c("STUDYID", "USUBJID", "SUBJID", "SITEID", "AGE", "SEX"),
        ADTTE = c(
          "STUDYID", "USUBJID", "SUBJID", "SITEID",
          "PARAM", "PARAMCD", "ARM", "ARMCD", "AVAL", "CNSR"
        )
      )
    ),
    # subsetting or changing order of datasets
    tm_data_table(
      label = "Datasets order",
      variables_selected = list(
        ADSL = c("STUDYID", "USUBJID", "SUBJID", "SITEID", "AGE", "SEX"),
        ADLB = c(
          "STUDYID", "USUBJID", "SUBJID", "SITEID",
          "PARAM", "PARAMCD", "AVISIT", "AVISITN", "AVAL", "CHG"
        )
      ),
      datasets_selected = c("ADTTE", "ADLB", "ADSL")
    ),
    # advanced usage of DT options and extensions
    tm_data_table(
      label = "Advanced DT usage",
      dt_args = list(extensions = c("Buttons", "ColReorder", "FixedHeader")),
      dt_options = list(
        searching = FALSE,
        pageLength = 30,
        lengthMenu = c(5, 15, 25, 50, 100),
        scrollX = FALSE,
        dom = "lBrtip",
        buttons = c("copy", "csv", "excel", "pdf", "print"),
        colReorder = TRUE,
        fixedHeader = TRUE
      )
    )
  )
)
```

## Run the app

A simple `shiny::shinyApp` call will let you run the app. 
Note that app is only displayed when running this code inside an R session.  

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```

