---
title: "Using scatterplot"
author: "Dawid Kałędkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Using scatterplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Teal application to use scatter plot with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside an scatter plot module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(dplyr) # used to modify data sets
```

## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADSL2` A wide data set with subject data
3. `ADRS` A long data set with response data for subjects at different time points of the study
4. `ADTTE` A long data set with time to event data
5. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
# nolint start
ADSL <- teal.modules.general::rADSL
ADSL2 <- teal.modules.general::rADSL %>%
  mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))
ADRS <- teal.modules.general::rADRS
ADTTE <- teal.modules.general::rADTTE
ADLB <- teal.modules.general::rADLB %>%
  mutate(CHGC = as.factor(case_when(
    CHG < 1 ~ "N",
    CHG > 1 ~ "P",
    TRUE ~ "-"
  )))
# nolint end
```

## Create an `app` variable

This is the most important section. We will use the [`teal::init`](https://insightsengineering.github.io/teal/reference/init) function to
create an app. The data will be handed over using [`teal.data::cdisc_data`](https://insightsengineering.github.io/teal.data/reference/cdisc_data). The app
itself will be constructed by multiple calls of `tm_g_scatterplot` using different
combinations of data sets.

```{r ggExtra, include = FALSE}
ggextra_available <- requireNamespace("ggExtra", quietly = TRUE)
```
```{r, include = !ggextra_available}
# NOTE: The code will not be run as package ggExtra is not installed.
```
```{r, eval = ggextra_available, echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- teal::init(
  data = teal.data::cdisc_data(
    teal.data::cdisc_dataset("ADSL", ADSL, code = "ADSL <- teal.modules.general::rADSL"),
    teal.data::cdisc_dataset(
      "ADSL2",
      ADSL2,
      keys = get_cdisc_keys("ADSL"),
      code = "ADSL2 <- teal.modules.general::rADSL %>%
              mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))"
    ),
    teal.data::cdisc_dataset("ADRS", ADRS, code = "ADRS <- teal.modules.general::rADRS"),
    teal.data::cdisc_dataset("ADTTE", ADTTE, code = "ADTTE <- teal.modules.general::rADTTE"),
    teal.data::cdisc_dataset("ADLB", ADLB,
      code = "ADLB <- teal.modules.general::rADLB %>%
            mutate(CHGC = as.factor(case_when(
            CHG < 1 ~ 'N',
            CHG > 1 ~ 'P',
            TRUE ~ '-'
            )))"
    ),
    check = TRUE
  ),
  modules = teal::modules(
    modules(
      label = "Scatterplot",
      tm_g_scatterplot(
        label = "Single wide dataset",
        x = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("AGE", "BMRKR1", "BMRKR2")),
            selected = "AGE",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("AGE", "BMRKR1", "BMRKR2")),
            selected = "BMRKR1",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL, c("RACE", "SEX")),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE
          )
        )
      ),
      # .. Two wide -----
      tm_g_scatterplot(
        label = "Two wide datasets",
        x = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("BMRKR1", "BMRKR2")),
            selected = "BMRKR1",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADSL2",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL2, c("AGE", "SEX")),
            selected = "AGE",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADSL2",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL2, c("COUNTRY", "AGE", "RACE")),
            selected = "COUNTRY",
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),
      tm_g_scatterplot(
        label = "Different long datasets",
        x = teal.transform::data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADRS),
            selected = "AVAL",
            multiple = FALSE,
            fixed = FALSE
          ),
          filter = teal.transform::filter_spec(
            label = "Select endpoint:",
            vars = c("PARAMCD", "AVISIT"),
            choices = value_choices(ADRS, c("PARAMCD", "AVISIT"), c("PARAM", "AVISIT")),
            selected = "OVRINV - SCREENING",
            multiple = FALSE
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADTTE",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADTTE),
            selected = "AVAL",
            multiple = FALSE,
            fixed = FALSE
          ),
          filter = teal.transform::filter_spec(
            label = "Select parameters:",
            vars = c("PARAMCD"),
            choices = value_choices(ADTTE, "PARAMCD", "PARAM"),
            selected = "OS",
            multiple = TRUE
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("AGE", "SEX")),
            selected = "AGE",
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),
      tm_g_scatterplot(
        label = "Wide and long datasets",
        x = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("SEX", "AGE", "BMRKR1", "COUNTRY")),
            selected = "AGE",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select measurement:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            label = "Selected variable:",
            choices = "AVAL",
            selected = "AVAL",
            multiple = FALSE,
            fixed = TRUE
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("SEX", "AGE", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),
      tm_g_scatterplot(
        label = "Same long datasets (same subsets)",
        x = teal.transform::data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AVAL", "BMRKR1", "BMRKR2")),
            selected = "AVAL",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AVAL", "BMRKR1", "BMRKR2")),
            selected = "BMRKR1",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AGE", "SEX", "RACE")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        )
      ),
      tm_g_scatterplot(
        label = "Same long datasets (different subsets)",
        x = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = "AVAL",
            selected = "AVAL",
            multiple = FALSE,
            fixed = TRUE
          )
        ),
        y = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = "AVAL",
            selected = "AVAL",
            multiple = FALSE,
            fixed = TRUE
          )
        ),
        color_by = teal.transform::data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = variable_choices(ADLB, c("RACE", "SEX")),
            selected = "SEX",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        )
      )
    )
  )
)
```

## Run the app

A simple `shiny::shinyApp` call will let you run the app.
Note that app is only displayed when running this code inside an R session.

```{r, eval = ggextra_available, echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```
