---
title: "Using outliers module"
author: "Mahmoud Hallal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Using outliers module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Teal application to analyze and report outliers with various datasets types.

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside the outliers module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(dplyr) # used to modify data sets
```

## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADRS` A long data set with response data for subjects at different time points of the study
3. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
# nolint start
ADSL <- teal.modules.general::rADSL
ADRS <- teal.modules.general::rADRS
ADLB <- teal.modules.general::rADLB
# nolint end
```

## Create an `app` variable

This is the most important section. We will use the [`teal::init`](https://insightsengineering.github.io/teal/reference/init) function to
create an app. The data will be handed over using [`teal.data::cdisc_data`](https://insightsengineering.github.io/teal.data/reference/cdisc_data). The app
itself will be constructed by multiple calls of `tm_outliers` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- teal::init(
  data = teal.data::cdisc_data(
    teal.data::cdisc_dataset("ADSL", ADSL, code = "ADSL <- teal.modules.general::rADSL"),
    teal.data::cdisc_dataset("ADRS", ADRS, code = "ADRS <- teal.modules.general::rADRS"),
    teal.data::cdisc_dataset("ADLB", ADLB, code = "ADLB <- teal.modules.general::rADLB"),
    check = TRUE
  ),
  modules = teal::modules(
    # tm_outliers ----
    modules(
      label = "Outliers module",
      tm_outliers(
        label = "Single wide dataset",
        outlier_var = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("AGE", "BMRKR1")),
            selected = "AGE",
            fixed = FALSE
          )
        ),
        categorical_var = teal.transform::data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL, subset = names(Filter(isTRUE, sapply(ADSL, is.factor)))),
            selected = "RACE",
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),
      tm_outliers(
        label = "Wide and long datasets",
        outlier_var = list(
          teal.transform::data_extract_spec(
            dataname = "ADSL",
            select = select_spec(
              label = "Select variable:",
              choices = variable_choices(ADSL, c("AGE", "BMRKR1")),
              selected = "AGE",
              fixed = FALSE
            )
          ),
          teal.transform::data_extract_spec(
            dataname = "ADLB",
            select = select_spec(
              label = "Select variable:",
              choices = variable_choices(ADLB, c("AVAL", "CHG2")),
              selected = "AVAL",
              multiple = FALSE,
              fixed = FALSE
            )
          )
        ),
        categorical_var =
          teal.transform::data_extract_spec(
            dataname = "ADSL",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADSL, subset = names(Filter(isTRUE, sapply(ADSL, is.factor)))),
              selected = "RACE",
              multiple = FALSE,
              fixed = FALSE
            )
          )
      ),
      tm_outliers(
        label = "Multiple long datasets",
        outlier_var = list(
          teal.transform::data_extract_spec(
            dataname = "ADRS",
            select = select_spec(
              label = "Select variable:",
              choices = variable_choices(ADRS, c("ADY", "EOSDY")),
              selected = "ADY",
              fixed = FALSE
            )
          ),
          teal.transform::data_extract_spec(
            dataname = "ADLB",
            select = select_spec(
              label = "Select variable:",
              choices = variable_choices(ADLB, c("AVAL", "CHG2")),
              selected = "AVAL",
              multiple = FALSE,
              fixed = FALSE
            )
          )
        ),
        categorical_var = list(
          teal.transform::data_extract_spec(
            dataname = "ADRS",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADRS, c("ARM", "ACTARM")),
              selected = "ARM",
              multiple = FALSE,
              fixed = FALSE
            )
          ),
          teal.transform::data_extract_spec(
            dataname = "ADLB",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADLB, subset = names(Filter(isTRUE, sapply(ADLB, is.factor)))),
              selected = "RACE",
              multiple = FALSE,
              fixed = FALSE
            )
          )
        )
      )
    )
  )
)
```
## Run the app

A simple `shiny::shinyApp` call will let you run the app. 
Note that app is only displayed when running this code inside an R session.  

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```
