---
title: "Using bivariate plot"
author: "Dawid Kałędkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Using bivariate plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Teal application to use bivariate plots with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside an bivariate plot module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App



## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(scda) # used to create data sets
library(dplyr) # used to modify data sets
```


## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADSL2` A wide data set with subject data
3. `ADRS` A long data set with response data for subjects at different time points of the study
4. `ADTTE` A long data set with time to event data
5. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
ADSL <- synthetic_cdisc_data("latest")$adsl #nolint
ADSL2 <- synthetic_cdisc_data("latest")$adsl %>% #nolint
  mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))
ADRS <- synthetic_cdisc_data("latest")$adrs #nolint
ADTTE <- synthetic_cdisc_data("latest")$adtte #nolint
ADLB <- synthetic_cdisc_data("latest")$adlb %>% #nolint
  mutate(CHGC = as.factor(case_when(
    CHG < 1 ~ "N",
    CHG > 1 ~ "P",
    TRUE ~ "-"
  )))
```

## Create an `app` variable

This is the most important section. We will use the `teal::init` function to
create an app. The data will be handed over using `teal::cdisc_data`. The app
itself will be constructed by multiple calls of `tm_g_bivariate` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- init(
  data = cdisc_data(
    cdisc_dataset("ADSL", ADSL, code = "ADSL <- synthetic_cdisc_data(\"latest\")$adsl"),
    dataset("ADSL2", ADSL2, keys = get_cdisc_keys("ADSL"),
      code = "ADSL2 <- synthetic_cdisc_data(\"latest\")$adsl %>%
              mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))"),
    cdisc_dataset("ADRS", ADRS, code = "ADRS <- synthetic_cdisc_data(\"latest\")$adrs"),
    cdisc_dataset("ADTTE", ADTTE, code = "ADTTE <- synthetic_cdisc_data(\"latest\")$adtte"),
    cdisc_dataset("ADLB", ADLB,
    code = "ADLB <- synthetic_cdisc_data(\"latest\")$adlb %>%
            mutate(CHGC = as.factor(case_when(
            CHG < 1 ~ 'N',
            CHG > 1 ~ 'P',
            TRUE ~ '-'
            )))"),
    check = TRUE
  ),
  modules = root_modules(
    # tm_g_bivariate ------
    modules(
      label = "Bivariate plot",
      tm_g_bivariate(
        label = "Single wide dataset",
        x = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = "BMRKR1",
            fixed = FALSE
          )
        ),
        y = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = "SEX",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE
          )
        )
      ),

      tm_g_bivariate(
        label = "Two wide datasets",
        x = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL, c("BMRKR1", "AGE", "SEX", "STRATA1", "RACE")),
            selected = c("BMRKR1"),
            multiple = FALSE
          )
        ),
        y = data_extract_spec(
          dataname = "ADSL2",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL2, c("COUNTRY", "AGE", "RACE")),
            selected = "RACE",
            multiple = FALSE
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADSL2",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL2),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE
          )
        )
      ),

      tm_g_bivariate(
        label = "Multiple different long datasets",
        x = data_extract_spec(
          dataname = "ADRS",
          filter = filter_spec(
            label = "Select endpoints:",
            vars = c("PARAMCD", "AVISIT"),
            choices = value_choices(ADRS, c("PARAMCD", "AVISIT"), c("PARAM", "AVISIT")),
            selected = "OVRINV - END OF INDUCTION",
            multiple = TRUE
          ),
          select = select_spec(
            choices = variable_choices(ADRS, c("AVALC", "AVAL")),
            selected = "AVALC",
            multiple = FALSE
          )
        ),
        y = data_extract_spec(
          dataname = "ADTTE",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADTTE, c("AVAL", "CNSR")),
            selected = "AVAL",
            multiple = FALSE,
            fixed = FALSE
          ),
          filter = filter_spec(
            label = "Select endpoint:",
            vars = c("PARAMCD"),
            choices = value_choices(ADTTE, "PARAMCD", "PARAM"),
            selected = "OS",
            multiple = FALSE
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADRS",
          filter = filter_spec(
            label = "Select endpoints:",
            vars = c("PARAMCD", "AVISIT"),
            choices = value_choices(ADRS, c("PARAMCD", "AVISIT"), c("PARAM", "AVISIT")),
            selected = "OVRINV - SCREENING",
            multiple = TRUE
          ),
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADRS, c("SEX", "RACE", "COUNTRY", "ARM", "PARAMCD", "AVISIT")),
            selected = "SEX",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL, c("SEX", "RACE")),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE
          )
        ),
        color_settings = TRUE,
        color = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        fill = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        size = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("AGE", "BMRKR1")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        plot_height = c(600, 200, 2000),
        ggtheme = "gray"
      ),
      tm_g_bivariate(
        label = "Wide and long datasets",
        x = data_extract_spec(
          dataname = "ADRS",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADRS, "PARAMCD", "PARAM"),
              selected = levels(ADRS$PARAMCD)[1],
              multiple = FALSE,
              label = "Select response:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADRS$AVISIT),
              selected = levels(ADRS$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = variable_choices(ADRS, c("AVALC", "AVAL")),
            selected = "AVALC",
            multiple = FALSE,
            label = "Select variable:"
          )
        ),
        y = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("BMRKR1", "SEX", "AGE", "RACE", "COUNTRY")),
            selected = "BMRKR1",
            multiple = FALSE,
            label = "Select variable:",
            fixed = FALSE
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("SEX", "RACE", "ARMCD", "PARAMCD")),
            selected = "SEX",
            multiple = FALSE,
            label = "Select variable:"
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("SEX", "RACE", "ARMCD", "PARAMCD", "AVISIT")),
            selected = "ARMCD",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        )
      ),

      tm_g_bivariate(
        label = "Wide and multiple long datasets",
        x = data_extract_spec(
          dataname = "ADRS",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADRS, "PARAMCD", "PARAM"),
              selected = levels(ADRS$PARAMCD)[1],
              multiple = FALSE,
              label = "Select response:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADRS$AVISIT),
              selected = levels(ADRS$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = variable_choices(ADRS, c("AVALC", "AVAL")),
            selected = "AVALC",
            multiple = FALSE,
            label = "Select variable:"
          )
        ),
        y = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("BMRKR1", "SEX", "AGE", "RACE", "COUNTRY")),
            selected = "BMRKR1",
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select measurement:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = "ARMCD",
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "AGE", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        color_settings = TRUE,
        color = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        fill = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        size = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("AGE", "BMRKR1")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        plot_height = c(600, 200, 2000),
        ggtheme = "gray"
      ),

      tm_g_bivariate(
        label = "Same long datasets (same subset)",
        x = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AVALC", "AVAL")),
            selected = "AVALC",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        y = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("SEX", "RACE", "COUNTRY", "ARMCD", "BMRKR1", "BMRKR2")),
            selected = "BMRKR1",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        row_facet = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AVISIT", "PARAMCD")),
            selected = "PARAMCD",
            multiple = TRUE,
            label = "Select variables:"
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADRS",
          select = select_spec(
            choices = variable_choices(ADRS, c("AVISIT", "PARAMCD")),
            selected = "AVISIT",
            multiple = TRUE,
            label = "Select variables:"
          )
        )
      ),

      tm_g_bivariate(
        label = "Same datasets (different subsets)",
        x = data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[2],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = "AVAL",
            selected = "AVAL",
            multiple = FALSE,
            fixed = TRUE
          )
        ),
        y = data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select visit:"
            )
          ),
          select = select_spec(
            choices = "AVAL",
            selected = "AVAL",
            multiple = FALSE,
            fixed = TRUE
          )
        ),
        use_density = FALSE,
        row_facet = data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[1],
              multiple = FALSE,
              label = "Select category:"
            )
          ),
          select = select_spec(
            choices = variable_choices(ADLB, c("RACE", "SEX", "ARMCD", "ACTARMCD")),
            selected = NULL,
            multiple = TRUE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        col_facet = data_extract_spec(
          dataname = "ADLB",
          filter = list(
            filter_spec(
              vars = "PARAMCD",
              choices = value_choices(ADLB, "PARAMCD", "PARAM"),
              selected = levels(ADLB$PARAMCD)[1],
              multiple = FALSE,
              label = "Select lab:"
            ),
            filter_spec(
              vars = "AVISIT",
              choices = levels(ADLB$AVISIT),
              selected = levels(ADLB$AVISIT)[3],
              multiple = FALSE,
              label = "Select category:"
            )
          ),
          select = select_spec(
            choices = variable_choices(ADLB, c("RACE", "SEX", "ARMCD", "ACTARMCD")),
            selected = "ARMCD",
            multiple = TRUE,
            fixed = FALSE,
            label = "Select variables:"
          )
        ),
        color_settings = TRUE,
        color = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        fill = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("SEX", "RACE", "COUNTRY")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        size = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            choices = variable_choices(ADSL, c("AGE", "BMRKR1")),
            selected = NULL,
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        plot_height = c(600, 200, 2000),
        ggtheme = "gray"
      )
    )
  )
)
```


## Run the app

A simple `shiny::shinyApp` call will let you run the app.
Note that app is only displayed when running this code inside an R session.

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```
