---
title: "Using cross table"
author: "Dawid Kałędkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Cross table app construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Teal application to use cross table with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside an cross table module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(random.cdisc.data) # used to create data sets
library(dplyr) # used to modify data sets
```

## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADSL_2` A wide data set with subject data
3. `ADRS` A long data set with response data for subjects at different time points of the study
4. `ADTTE` A long data set with time to event data
5. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
ADSL <- radsl(cached = TRUE)
ADSL_2 <- radsl(cached = TRUE)
ADRS <- radrs(cached = TRUE)
ADTTE <- radtte(cached = TRUE)
ADLB <- radlb(cached = TRUE) %>%
  mutate(CHGC = as.factor(case_when(
    CHG < 1 ~ "N",
    CHG > 1 ~ "P",
    TRUE ~ "-"
  )))
```

## Create an `app` variable

This is the most important section. We will use the `teal::init` function to
create an app. The data will be handed over using `teal::cdisc_data`. The app
itself will be constructed by multiple calls of `tm_t_crosstable` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- init(
  data = cdisc_data(
    cdisc_dataset("ADSL", ADSL),
    dataset("ADSL_2", ADSL_2, keys = get_cdisc_keys("ADSL")),
    cdisc_dataset("ADRS", ADRS),
    cdisc_dataset("ADTTE", ADTTE),
    cdisc_dataset("ADLB", ADLB),
    code = "ADSL <- radsl(cached = TRUE)
            ADSL_2 <- radsl(cached = TRUE)
            ADRS <- radrs(cached = TRUE)
            ADTTE <- radtte(cached = TRUE)
            ADLB <- radlb(cached = TRUE) %>%
            mutate(CHGC = as.factor(case_when(
            CHG < 1 ~ 'N',
            CHG > 1 ~ 'P',
            TRUE ~ '-'
            )))",
    check = TRUE
  ),
  modules = root_modules(
    modules(
      label = "Cross table",
      tm_t_crosstable(
        label = "Single wide dataset",
        x = data_extract_spec(
          "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = names(ADSL)[5],
            multiple = FALSE,
            fixed = FALSE
          )
        ),
        y = data_extract_spec(
          "ADSL",
          select = select_spec(
            label = "Select variable:",
            choices = variable_choices(ADSL),
            selected = names(ADSL)[6],
            multiple = FALSE,
            fixed = FALSE
          )
        )
      ),

      tm_t_crosstable(
        label = "Same long datasets (different subsets)",
        x = data_extract_spec(
          dataname = "ADLB",
          filter = filter_spec(
            vars = "PARAMCD",
            choices = value_choices(ADLB, "PARAMCD", "PARAM"),
            selected = levels(ADLB$PARAMCD)[1],
            multiple = FALSE
          ),
          select = select_spec(
            choices = variable_choices(ADLB),
            selected = "AVISIT",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        ),
        y = data_extract_spec(
          dataname = "ADLB",
          filter = filter_spec(
            vars = "PARAMCD",
            choices = value_choices(ADLB, "PARAMCD", "PARAM"),
            selected = levels(ADLB$PARAMCD)[1],
            multiple = FALSE
          ),
          select = select_spec(
            choices = variable_choices(ADLB),
            selected = "LOQFL",
            multiple = FALSE,
            fixed = FALSE,
            label = "Select variable:"
          )
        )
      )
    )
  )
)
```

## Run the app

A simple `shiny::shinyApp` call will let you run the app. 

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```
