---
title: "Using scatterplot matrix"
author: "Dawid Kałędkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
runtime: shiny
vignette: >
  %\VignetteIndexEntry{Scatterplot matrix app construction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


Teal application to use scatter plot matrix with various datasets types

This vignette will guide you through 4 parts to create a teal application using
various types of datasets inside an scatter plot matrix module:

1. Load Libraries
2. Create data sets
3. Create an `app` variable
4. Run the App

## Loading libraries

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
library(teal.modules.general) # used to create the app
library(random.cdisc.data) # used to create data sets
library(dplyr) # used to modify data sets
```

## Create data sets

Inside this app 5 datasets will be used

1. `ADSL` A wide data set with subject data
2. `ADSL2` A wide data set with subject data
3. `ADRS` A long data set with response data for subjects at different time points of the study
4. `ADTTE` A long data set with time to event data
5. `ADLB` A long data set with lab measurements for each subject

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
ADSL <- radsl(cached = TRUE)
ADSL2 <- radsl(cached = TRUE) %>% 
  mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))
ADRS <- radrs(cached = TRUE)
ADTTE <- radtte(cached = TRUE)
ADLB <- radlb(cached = TRUE) %>%
  mutate(CHGC = as.factor(case_when(
    CHG < 1 ~ "N",
    CHG > 1 ~ "P",
    TRUE ~ "-"
  )))
```

## Create an `app` variable

This is the most important section. We will use the `teal::init` function to
create an app. The data will be handed over using `teal::cdisc_data`. The app
itself will be constructed by multiple calls of `tm_g_scatterplotmatrix` using different
combinations of data sets.

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hide"}
app <- init(
  data = cdisc_data(
    cdisc_dataset("ADSL", ADSL),
    dataset("ADSL2", ADSL2, keys = get_cdisc_keys("ADSL")),
    cdisc_dataset("ADRS", ADRS),
    cdisc_dataset("ADTTE", ADTTE),
    cdisc_dataset("ADLB", ADLB),
    code = "ADSL <- radsl(cached = TRUE)
            ADSL2 <- radsl(cached = TRUE) %>% 
            mutate(TRTDUR = round(as.numeric(TRTEDTM - TRTSDTM), 1))
            ADRS <- radrs(cached = TRUE)
            ADTTE <- radtte(cached = TRUE)
            ADLB <- radlb(cached = TRUE) %>%
            mutate(CHGC = as.factor(case_when(
            CHG < 1 ~ 'N',
            CHG > 1 ~ 'P',
            TRUE ~ '-'
            )))",
    check = TRUE
  ),
  modules = root_modules(
    modules(
      label = "Scatterplot matrix",
      # .. single wide ----
      tm_g_scatterplotmatrix(
        label = "Single wide dataset",
        variables = data_extract_spec(
          dataname = "ADSL",
          select = select_spec(
            label = "Select variables:",
            choices = variable_choices(ADSL),
            selected = c("AGE", "RACE", "SEX", "BMRKR1", "BMRKR2"),
            multiple = TRUE,
            fixed = FALSE
          )
        )
      ),

      tm_g_scatterplotmatrix(
        label = "Multiple wide datasets",
        variables = list(
          data_extract_spec(
            dataname = "ADSL",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADSL),
              selected = c("AGE", "ACTARM", "SEX", "BMRKR1"),
              multiple = TRUE,
              fixed = FALSE
            )
          ),
          data_extract_spec(
            dataname = "ADSL2",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADSL2),
              selected = c("COUNTRY", "ACTARM", "STRATA1"),
              multiple = TRUE,
              fixed = FALSE
            )
          )
        )
      ),

      tm_g_scatterplotmatrix(
        "One long dataset",
        variables = data_extract_spec(
          dataname = "ADTTE",
          select = select_spec(
            choices = variable_choices(ADTTE, c("AVAL", "BMRKR1", "BMRKR2")),
            selected = c("AVAL", "BMRKR1", "BMRKR2"),
            multiple = TRUE,
            fixed = FALSE,
            label = "Select variables:"
          )
        )
      ),

      tm_g_scatterplotmatrix(
        label = "Two long datasets",
        variables = list(
          data_extract_spec(
            dataname = "ADRS",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADRS),
              selected = c("AVAL", "AVALC"),
              multiple = TRUE,
              fixed = FALSE
            ),
            filter = filter_spec(
              label = "Select endpoints:",
              vars = c("PARAMCD", "AVISIT"),
              choices = value_choices(ADRS, c("PARAMCD", "AVISIT"), c("PARAM", "AVISIT")),
              selected = "OVRINV - SCREENING",
              multiple = FALSE
            )
          ),
          data_extract_spec(
            dataname = "ADTTE",
            select = select_spec(
              label = "Select variables:",
              choices = variable_choices(ADTTE),
              selected = c("AVAL", "CNSR"),
              multiple = TRUE,
              fixed = FALSE
            ),
            filter = filter_spec(
              label = "Select parameters:",
              vars = "PARAMCD",
              choices = value_choices(ADTTE, "PARAMCD", "PARAM"),
              selected = "OS",
              multiple = TRUE
            )
          )
        )
      )
    )
  )
)

```

## Run the app

A simple `shiny::shinyApp` call will let you run the app. 

```{r echo=TRUE}
shinyApp(app$ui, app$server, options = list(height = 1024, width = 1024))
```

